---
title: "Analysis of Workplace-Themed Term Co-Occurrence Matrices"
author: "Ben Kubacki"
date: "13 December 2024"
output: 
  github_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, include=TRUE, comment="")

library(tidyverse)
library(readxl)

# web scraping
library(rvest)

# tokenizer
library(tidytext)

# word embeddings
library(text2vec)
library(Rling)
library(cluster)
library(tm)
```

# Read in data

> Data is already lemmatized and POS-tagged.

```{r}

# c("data_private/wlp_newspaper_lsp/wlp_news_2012.txt") %>% 
#   read_tsv(col_names = F)
  
  news_wlp <- c("data_private/wlp_newspaper_lsp/wlp_news_1990.txt",
           "data_private/wlp_newspaper_lsp/wlp_news_1991.txt",
           "data_private/wlp_newspaper_lsp/wlp_news_1992.txt",
           "data_private/wlp_newspaper_lsp/wlp_news_1993.txt",
           "data_private/wlp_newspaper_lsp/wlp_news_1994.txt",
           "data_private/wlp_newspaper_lsp/wlp_news_1995.txt",
           "data_private/wlp_newspaper_lsp/wlp_news_1996.txt",
           "data_private/wlp_newspaper_lsp/wlp_news_1997.txt",
           "data_private/wlp_newspaper_lsp/wlp_news_1998.txt",
           "data_private/wlp_newspaper_lsp/wlp_news_1999.txt",
           "data_private/wlp_newspaper_lsp/wlp_news_2000.txt",
           "data_private/wlp_newspaper_lsp/wlp_news_2001.txt",
           "data_private/wlp_newspaper_lsp/wlp_news_2002.txt",
           "data_private/wlp_newspaper_lsp/wlp_news_2003.txt",
           "data_private/wlp_newspaper_lsp/wlp_news_2004.txt",
           "data_private/wlp_newspaper_lsp/wlp_news_2005.txt",
           "data_private/wlp_newspaper_lsp/wlp_news_2006.txt",
           "data_private/wlp_newspaper_lsp/wlp_news_2007.txt",
           "data_private/wlp_newspaper_lsp/wlp_news_2008.txt",
           "data_private/wlp_newspaper_lsp/wlp_news_2009.txt",
           "data_private/wlp_newspaper_lsp/wlp_news_2010.txt",
           "data_private/wlp_newspaper_lsp/wlp_news_2011.txt",
           "data_private/wlp_newspaper_lsp/wlp_news_2012.txt") %>% 
  read_tsv(col_names = F)

```
# Clean the data

## Tidy columns

- Rename "word", "lemma", "pos"  
- Add document_id column  
- Fill document_id column NAs with document_id

```{r}

## Rename columns
news_wlp <- news_wlp %>%
  select(word = X1, lemma = X2, pos = X3)

## Add column marked by document number (when lemma & pos are NA), keep the rest NA
news_wlp <- news_wlp %>%
  mutate(document_id = ifelse(is.na(news_wlp$pos) & is.na(news_wlp$lemma), str_sub(word, 3,9),NA))

## Replace NAs with document_id 
news_wlp <- news_wlp %>%
  fill(document_id)

```

## Filter POS to include only adjectives, nouns, and verbs. 

> Compare full list against included list in code chunk

POS Tags [reference list](https://ucrel.lancs.ac.uk/claws7tags.html)

- Some data is marked with multiple POS (i.e., jj_nn)  
- Create regex separated by `|` to filter for all cases, including multiples

```{r}

pos_filtered <- str_to_lower(c("JJ", "JJR", "JJT", "JK", "NN", "NN1", "NN2", "NNA", "NNB", "NNL1", "NNL2", "NNO", "NNO2", "NNT1", "NNT2", "NNU", "NNU1", "NNU2", "VV0", "VVD", "VVG", "VVGK", "VVI", "VVN", "VVNK", "VVZ"))

pos_filtered_c <- str_c(pos_filtered, collapse="|")

news_wlp <- news_wlp %>% 
  filter(str_detect(pos, pos_filtered_c))

```

# Build TCM Matrix

Source of demo, including object names, code, and comments inside code: (Selivanov, 2023)[https://cran.r-project.org/web/packages/text2vec/vignettes/glove.html]

## Create tokens

- Remove punctuation and contractions (may already be removed from POS filtering)

```{r}

## Create tokens object

tokens <- news_wlp %>% select(document_id, lemma) %>% chop(lemma) %>% pull(lemma, document_id) #%>% 
  #as.list()

stop_punctuation_regex <- "[<p>]|[\\,\\.\\:\\;\\*\\@\\'\\\"\\-\\$\\(\\)\\?]|(\\-\\-)|('s)|('d)|('t)|(n't)|('ve)|('re)|('ll)"

tokens <- map(tokens, ~ str_to_lower(str_subset(.x, stop_punctuation_regex, negate=T)))

```

## Create iterator

```{r}

# Create vocabulary. Terms will be unigrams (simple words).
it <- itoken(tokens, progressbar = FALSE)

```

## Create and filter vocabulary for infrequent terms and stop-words

### Create stop-word list

```{r}

## Make table with tidytext "snowball" and tm "en" and my own additions
  
# tidytext
stop_words_snowball <- filter(stop_words, lexicon=="snowball") %>% select(word)

# tm
stopwords(kind = "en")

# My additions
my_stopwords <- c("also", "even", "one", "yet", "y")

# Aggregated
stop_word_list <- c(stopwords(kind="en"), stop_words_snowball$word, my_stopwords)

```

### Create vocabulary

```{r}
vocab <- create_vocabulary(it)

# Remove very infrequent terms using min term count =2

## Determine which max count you want to remove if any
# vocab %>% arrange(desc(term_count)) %>% head(100)

# add argument term_count_max = 170000L)
vocab <- prune_vocabulary(vocab, term_count_min = 2L)

vocab <- vocab %>% filter(!term %in% stop_word_list)
```

## Create vectorizer and term co-occurrence matrix

```{r}
vectorizer <- vocab_vectorizer(vocab)

tcm <- create_tcm(it, vectorizer, skip_grams_window = 5L)

```

## Create vectors

```{r}

glove <-  GlobalVectors$new(rank = 50, x_max = 10)

wv_main <-  glove$fit_transform(tcm, n_iter = 10, convergence_tol = 0.01, n_threads = 8)

wv_context = glove$components

word_vectors = wv_main + t(wv_context)

```


Cosine similarities per target word: 

"work" should include both nouns and verbs, unfortunately

```{r}

## Target word "work"
cos_work <- sim2(x = word_vectors, y = word_vectors["work", , drop=FALSE], method = "cosine", norm = "l2")[,0:1] %>% sort(decreasing = T)

## Target word "meeting"
cos_meeting <- sim2(x = word_vectors, y = word_vectors["meeting", , drop=FALSE], method = "cosine", norm = "l2")[,0:1] %>% sort(decreasing = T)

## Target word "job"
cos_job <- sim2(x = word_vectors, y = word_vectors["job", , drop=FALSE], method = "cosine", norm = "l2")[,0:1] %>% sort(decreasing = T)

## Target word "interview"

## Target word "review"


```


```{r}

cos_tbl_work <- as_tibble(cos_work, rownames="words") %>% rename(cos_sim = value)

## Consider filtering by cos_sim
filter(cos_sim < 1 | >= 0.3)

## Add target word to facilitate join
cos_tbl_work <- cos_tbl_work %>%
  mutate(target = "work")

## Join the target word tibbles
cos_tbl <- full_join(cos_tbl_work, cos_tbl_meeting)


```

Write CSV

```{r}

cos_tbl %>% slice_sample() %>% write_csv()

```

Visualize


```{r}
arrange(proportion)

ggplot(slice(cos_tbl, 1:50), aes(x = words, y = cos_sim)) + geom_col()

```


```{r}

# read_csv("data_private/test_cos_25k.csv")

data_work <- read_csv("data_private/cosine_work.csv")

data_work %>% head(20)

data_work %>% summary()

data_work %>% 
  head(21) %>% 
  ggplot(aes(words, cos_sim))+
  geom_point(alpha = 0)+
  geom_text(aes(label = words))


cos_tbl_25k %>% 
  group_by(target) %>% 
  slice(1:21) %>% 
  ggplot(aes(words, cos_sim))+
  geom_point(alpha = 0)+
  geom_text(aes(label = words, colour = target))
  
```


```{r fig.align="center", echo = FALSE,fig.width = 14}
cos_tbl_25k %>% 
  group_by(target) %>% 
  slice(1:21) %>% 
  ggplot(aes(words, cos_sim))+
  geom_point(alpha = 0)+
  geom_text(aes(label = words, colour = target))

data_work %>% 
  head(50) %>% 
  ggplot(aes(words, cos_sim))+
  geom_point(alpha = 0)+
  geom_text(aes(label = words))
```

# Session info

```{r}
sessionInfo()
```
